Spatial population structure drives extinction dynamics in climate-induced range shifts

### 2018-09-06 Manuscript/Main.tex
Final update to the manuscript before the initial submission to Ecology

Instead, the stable distribution of dispersal phenotypes before the beginning of climate change played a large role in determining the ultimate fate of simulated populations. Simulated populations evolved a range of dispersal phenotypes over the $2000$ generations of stable climatic conditions in response to the potential for local adaptation and the severity of the habitat gradient at the range edge (Fig. S8-S10). The populations that survived climate change were those that happened to make up the upper end of these distributions, containing individuals with high dispersal phenotypes (Fig. 3c\&d). Comparing the dispersal phenotypes making up successful populations across different parameter combinations revealed a common threshold separating individuals from successful and extinct populations at each speed of climate change (Fig. S8-S10). In fact, this threshold corresponds to the dispersal phenotype necessary to produce an expansion wave traveling at exactly the speed of climate change. To show this, we used the well known approximation for the speed of an expanding population, $2\sqrt{rD}$~\citep{hastings2005spatial}, where $r$ is the intrinsic growth rate of the population and $D$ is the diffusion coefficient. The $2$ dimensional diffusion coefficient can be calculated directly from the dispersal phenotypes as $D=\frac{1}{4}d^{2}$ since $d^{2}$ represents the mean squared displacement of an individual with phenotype $d$ during a dispersal period (see the model description in the supplemental materials). Using these relationships and solving for the dispersal phenotype producing each of the speeds of climate change used in our simulations yields the threshold value indicated on Figures S8-S10 which distinguishes those populations that ultimately survived climate change from those that went extinct.

>>>>

Instead, the stable distribution of dispersal phenotypes prior to the onset of climate change played a key role in determining the ultimate fate of simulated populations. Simulated populations evolved a range of dispersal phenotypes over the $2000$ generations of stable climatic conditions in response to the potential for local adaptation and the severity of the habitat gradient at the range edge (Fig. S8-S10). The populations that survived climate change were those with higher initial dispersal phenotypes, occupying the right-tail of the distribution of observed dispersal phenotypes (Fig. 3c\&d). Comparing the dispersal phenotypes of individuals in successful populations across different parameter combinations revealed a common threshold separating individuals from successful and extinct populations at each speed of climate change (Fig. S8-S10). In fact, this threshold corresponded to the dispersal phenotype necessary to produce an expansion wave traveling at exactly the speed of climate change. To prove this, we used the well-known approximation for the speed of an expanding population, $2\sqrt{rD}$~\citep{hastings2005spatial}, where $r$ is the intrinsic growth rate of the population and $D$ is the diffusion coefficient. The $2$ dimensional diffusion coefficient can be calculated directly from the dispersal phenotypes as $D=\frac{1}{4}d^{2}$ since $d^{2}$ represents the mean squared displacement of an individual with phenotype $d$ during a dispersal period (see the model description in the supplemental materials). Using these relationships and solving for the dispersal phenotype producing each of the speeds of climate change used in our simulations yields the threshold value distinguishing those populations that ultimately survived climate change from those that went extinct (Figures S8-S10, vertical dashed line).
